### 2015-06-28 paper/supplementary.tex

%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
\documentclass{frontiers_suppmat} % for all articles

\usepackage{url}

\graphicspath{{../figures/}}

\def\articleType{Supplementary Material}
\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Bekolay {et~al.}}
\def\Authors{Trevor Bekolay\,$^{1,*}$,
  Terrence C. Stewart\,$^{1}$,
  Eric Hunsberger\,$^1$,
  Andrew Mundy\,$^2$ and
  Chris Eliasmith\,$^1$}
\def\Address{$^{1}$Centre for Theoretical Neuroscience,
  University of Waterloo,
  Waterloo, ON, Canada\\
  $^{2}$ Advanced Processors Technologies Research Group,
  School of Computer Science,
  University of Manchester,
  Manchester, UK}
\def\corrAuthor{Trevor Bekolay}
\def\corrAddress{Centre for Theoretical Neuroscience,
  David R. Cheriton School of Computer Science,
  University of Waterloo,
  200 University Avenue West,
  Waterloo, ON  N2L~3G1, Canada}
\def\corrEmail{tbekolay@uwaterloo.ca}

% \color{FrontiersColor} Is the color used in the Journal name, in the title, and the names of the sections.

\begin{document}
\onecolumn
\firstpage{1}

\title[Supplementary Material]{{\helveticaitalic{Supplementary Material}}:\\ \helvetica{ Benchmarking neuromorphic systems with Nengo}}
\author[\firstAuthorLast ]{\Authors}
\address{}
\correspondance{}
\extraAuth{}
\topic{Benchmarks and Challenges for Neuromorphic Engineering}

\maketitle

\section{Supplementary Tables and Figures}

\begin{figure}[!ht]
\begin{center}
  \includegraphics[width=85mm]{fig7}
\end{center}
\textbf{\refstepcounter{figure}\label{fig:01} Supplementary Figure \arabic{figure}.}
       {Run time for each model with probes removed on each backend.
         Each bar represents the mean run time across 100 instances
         of each model;
         error bars are 95\% confidence intervals.
         Run time is measured relative to real time;
         e.g., a value of two indicates that the model ran twice as slow
         as real time.}
\end{figure}

\end{document}