### 2015-08-29 paper/rebuttals.txt
Add reviews and rebuttal todo list

Todo list
=========

Work

- [ ] 1-4: Try to get unit tests running SpiNNaker
- [ ] 1-4: Why does removing probes slow down reference?
- [ ] 1-4: Ensure confidence intervals are real things
- [ ] 2-2: Find some citations of previous neuromorphic / simulator benchmarking
- [ ] 2-4: Collect information about current inputs?
- [ ] 2-5: Use different boxplot whisker type?

Writing

- [ ] 1-4: Rewrite lines 320-325
- [ ] 1-4: Write something about negative frequencies
- [ ] 1-4: Look up CPG frequencies or something
- [ ] 1-4: Explain compliance more (Table 1 especially)
- [ ] 1-5: Add discussion paragraph on efficiency
- [ ] 2-3: Add text on limitations to introduction / discussion
- [ ] 2-4: Add discussion on current input vs spike input?
- [ ] 2-4: Discuss why you'd want to set seeds
- [ ] 2-4: Comment on larger networks
- [ ] 2-4: Define fuzzy testing
- [ ] 2-4: Change table 1, 2 order
- [ ] 2-5: Better accuracy figure descriptions
- [ ] 2-5: Fig 7: Why are some slower with no probes
- [ ] 2-6: Fix typo
- [ ] 2-7: Fix citation for Stewart 2010

Reviewer 1
==========

1. What are the main findings reported in this manuscript?
----------------------------------------------------------

The main contribution of this work is a benchmarking suite designed to
test neuron simulators in a relevant and objective way. This is
performed by analyzing three fundamental metrics - compliance,
accuracy, and speed - under different network test
conditions. Compliance metrics determine feature coverage for a neural
simulator, while the accuracy and speed metrics are tested under four
neurally-relevant operations: representation with robustness to noise,
nonlinear computation, stability with recurrent dynamics, and
large-scale interaction of multiple recurrent populations, all
implemented with neurons. Finally, five neural simulators with
different algorithmic and hardware optimizations are tested on these
benchmarks.

2. Other comments on the introduction.
--------------------------------------

The introduction is well-written and motivates the study; the
background provides sufficient detail to understand the architecture
of this solution.

3. Other comments on the materials and methods.
-----------------------------------------------

The materials and methods describes well the metrics and models used
to evaluate the backends.

4. Other comments on the results.
---------------------------------

1. Further clarification should be given as to why the SpiNNaker
backend was not tested for compliance, as lines 276-77 are not
sufficient. The reason is not clear as build and run time for the
examined models in Figures 5-7 do not seem prohibitive to running 212
tests; additionally, SpiNNaker supports a wide variety of neuron
models and parameters, so it appears adequate to support the necessary
descriptions required by Nengo. Please clarify.

2. Lines 320-325 are not clear.

a) While Figure 6 does show an increase for the Nengo reference
backend run times by removing passthrough nodes, the run time appears
to be a 15% increase (~7x vs ~6x slower than real-time, Figure 6
green, BG Sequence to BG Sequence *), which is approximately the same
relative speedup as that offered by removing probes for the SpiNNaker
backend (~1.5x to ~1.2x, Figures 6 and 7, BG sequence, blue). So,
"significant" is a strong claim and should be clarified.

b) Additionally, Figure 7 seems to indicate that removing probes
actually makes the network slower for the reference backend for BG
sequence * (~8x vs ~6.8x, Figures 6 and 7, light blue, BG Sequence
*). Either outliers / standard deviations are stronger than they
appear from the plot, or the reason for the network slowing down with
probes removed should be explained.

c) Finally, the speedup by removing passthrough nodes for the OpenCL
backend (Figure 6, green, BG sequence to BG sequence*) disappears once
probes are removed (Figure 7, green, BG sequence to BG sequence*).

Essentially, please verify that the error bars on the plots are
representative of the true confidence intervals. If they are correct,
there seem to be some anomalies in the data (removing reporting output
slows down the reference backend for BG*, speedups in OpenCL become
slowdowns once reporting output is removed). If this all has to do
with build time / overhead / short runtimes / network latency, then
that should be mentioned in the text.

3. In Figure 3, the meaning of negative frequencies (-1 Hz, -2 Hz)
should be at least briefly mentioned. Presumably, since this is a
circular oscillator, it is being run in reverse. However, it's not
clear to me why negative frequencies should be at all different than
positive frequencies, and how the range -2 Hz to 2 Hz was chosen. A
citation that references this range in biology for CPGs, for example,
would bolster this choice.

4. Table 1, the compliance of the four backends, should be explained
more fully in the caption. The levels of coverage are surprisingly low
given that all backends are able to perform the main cognitive-level
tasks; common failure cases (e.g., missing neuron models) should be
mentioned to show why these are not necessary.

5. Other comments on the discussion or the conclusions.
-------------------------------------------------------

The initially stated conclusion that significant speedups can be
obtained with specialized hardware without sacrificing accuracy is
good, although perhaps not very surprising. However, the greater
challenge of creating a list of cognitively-relevant tasks and an
infrastructure for objective testing is significant. Moreover, the
improved accuracy of SpiNNaker with passthrough nodes removed,
discovered by these tests, is an encouraging sign that the
optimization loop of hardware and implementation through this test
framework can yield real improvements.

The biggest omission from this discussion and the manuscript is any
mention of efficiency. Neuromorphic systems have long held that their
primary advantage is significant efficiency, and the lack of ability
of the benchmarking to capture that is unfortunate. This could perhaps
be implemented as an optional query on the backend; the power
consumption of most CPUs and GPUs can be publicly referenced and
approximated, or the implementation could provide its own statistics
for specialized hardware. Perhaps the Brainstorm backend’s main
advantage is in power efficiency, and there should be resulting
statistics that can show that. Regardless of implementation, the
authors should at least mention that efficiency could be evaluated in
future work.

6. Please add here any further comments on this manuscript.
--------------------------------------------------------

This work addresses a very challenging topic - that is, how to
objectively compare systems designed for different objectives in a
high-dimensional benchmarking space. The approach of choosing
cognitive tasks and measuring the ability of neuromorphic systems to
perform them is insightful and relevant.

It does, however, represent only an initial step towards obtaining the
necessary concrete metrics to measure neuromorphic systems. Many
quantities remain untested in this framework; researchers who build
neuromorphic systems designed for neural simulation, the target
audience of this manuscript, focus on optimizing many of the following
quantities: simulation speed, simulation accuracy, energy per synaptic
event, accurate timing, number of neurons represented, number of
connections represented, repertoire of neural models, online learning,
and likely others. Of these, the test framework objectively compares
only along the first two, though the compliance tests likely would
stress some of the others. I hope to see future work along these lines
to expand the reach of these tests.

Nonetheless, this work provides a substantive proposal to begin
evaluating neuromorphic neural simulators.

Reviewer 2
==========

1. What are the main findings reported in this manuscript?
----------------------------------------------------------

This paper describes how the Nengo test suite can be used for
benchmarking neuromorphic systems and compare them to neural
simulators on general-purpose computers. It introduces the Nengo test
suite, describes three benchmark metrics (compliance, accuracy, speed)
and shows benchmark results for four tests on five different
backends. Of these backends, only one is actually running on
neuromorphic hardware, one is a simulator of a neuromorphic system,
and three are software simulators (with one employing GPUs). Benchmark
results are shortly discussed.

2. Does the introduction present the study in an appropriate context?
---------------------------------------------------------------------

No

The introduction gives a clear motivation for the work in the article,
but it does not state which works have been done previously on that
topic. For example, there are several articles investigating accuracy
in neuromorphic hardware for specific benchmark problems. As three of
the backends are software-based, it would also be appropriate here to
cite prior work on accuracy/speed in neural software simulators.

3. Other comments on the introduction.
--------------------------------------

As the Nengo test suite was not designed for benchmarking, it would be
helpful to have some indications on the limitations of the method (its
potential is described already pretty clear) - this could be done in
the Discussion section as well.

4. Other comments on the materials and methods.
-----------------------------------------------

I found the separation in Background and Methods helpful for
understanding, even if it does not follow the standard article
structure. Some points should be clarified:

- The benchmarks make use of current input to neurons. However, quite
  a lot of neuromorphic systems do only offer spike input, or current
  inputs are not so flexible to program. Of course, current input can
  be emulated by spike input, but that may add some additional
  noise/inaccuracy. This should be dicsussed in the Background/Methods
  sections; also, an indication on how the currents look like in the
  benchmark models would be helpful, so that designers of neuromorphic
  systems get an idea of what would be required.

- At several points, the authors mention that seeds can be set
  explicitely for deterministic experiment runs. Is there a reason for
  this? For benchmarking, setting a fixed seed is misleading in my
  opinion, as it suggests that all backends run exactly the same
  experiment. However, as the order in which the single parts of a
  network are generated is not the same, this is not the case in
  general. Furthermore, neuromorphic hardware systems may use built-in
  noise sources (i.e. true random numbers). All that requires a
  statistical comparison of results over several experiment
  repetitions, as the authors do in their benchmarks.

- The authors state that Nengo is well-suited for generating large
  networks. However, the tests they present are only limited in
  size. A comment on larger benchmark models in the test suite and/or
  the possibility to scale up the present benchmarks would be
  desirable.

Minor:

- I found the heading of section 2.2 ("Fuzzy testing") puzzling - is
  this a fixed term I am missing? In any case, it would be helpful to
  explain the term.

- Table 2 is referenced before Table 1 in the text - their order
  should be changed

5. Other comments on the results.
---------------------------------

I found the description of the Figures on accuracy too short. Even if
the accuracy results themselves are not the main focus of the article,
some more explanations for the results should be given. In particular,
in Figs. 1 and 2, why does the Nengo reference backend show a
significantly higher variability than the Brainstorm and distilled
backends? In Figs. 3 and 4, where do the deviations in the SpiNNaker
backend come from (my first guess would be the employed 1ms time step,
which is larger than the 0.1ms typically used in software simulators)?

Also, the whiskers in Figs. 1-4 seem to be only multiples of the box
sizes, so plotting them does not add any information - it may be even
misleading when not carefully reading the caption.

Some of the results in Fig. 7 are puzzling: For BG sequence*, the run
time for nengo and nengo_ocl is even higher without probes (compared
to Fig. 6), being well outside the 95% confidence intervals. This
should be explained.

6. Other comments on the discussion or the conclusions.
-------------------------------------------------------

As already mentioned for the Introduction, it would make the
Discussion section much more balanced if the limitations of the
approach would be discussed as well, together with some outlook to
potential improvements or complementary approaches.

Minor:

- typo in line 373: "We believe that THE these benchmarks..."

7. Please add here any further comments on this manuscript.
-----------------------------------------------------------

Overall, the article presents an attractive idea for benchmarking
neuromorphic systems. In my opinion its main weakness is the lack of
embedding in the current literature. Apart from that, it is a nice
contribution to the field, and the open source code is a good
opportunity for other neuromorphic system operators for benchmarking
their systems.

Minor:

- missing citation details for Stewart et al. 2010